---
title: "The SRP Package"
author: "Joseph T. Ornstein"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The SRP Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: refs.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

SRP is an `R` package that contains useful functions for implementing multilevel regression and poststratification (MRP) and stacked regression and poststratification (SRP).

## Motivation

Suppose we want to know how some public opinion varies by subnational unit. But we don't have surveys that were conducted at the unit-level, only a national-level survey. How can we use the information from the national survey to make inferences about the subnational level? This vignette walks through three techniques for doing so: disaggregation, multilevel regression and poststratification (MRP), and stacked regression and poststratification (SRP). It concludes with an introduction to synthetic poststratification.

## Installation

The `SRP` package is currently available on GitHub. You can install it using the `devtools` package. For this vignette, you'll also want to load the `tidyverse` package.

```{r installation, message = FALSE}
#devtools::install_github('joeornstein/SRP')
library(SRP)
library(tidyverse)
```

## The Data

The dataset (`trainset`) contains individual-level data on public opinion and demographic characteristics. It is simulated data, generated from the Monte Carlo in @Ornstein2019. It contains the following variables:


```{r data, message=FALSE}
trainset <- SRP::vignetteData

trainset
```

The variable $y$ is our outcome of interest, $x_1$ and $x_2$ are individual-level covariates, `unit` is the subnational unit ID, and `latitude`, `longitude`, and `unit_covariate` are characteristics of the subnational unit. 


## Disaggregation

Disaggregation is the most straightforward method to estimate. Simply take the unit-level means from the national survey. Note, however, that the number of observations within each unit is fairly small. As a result, disaggregation is unlikely to yield good estimates. This is why we adopt a model-based approach.

```{r disag, message=FALSE}
disag_estimates <- trainset %>% 
  group_by(unit) %>% 
  summarise(disag_estimate = mean(y),
            num = n())

disag_estimates
```


## MRP

Multilevel regression and poststratification (MRP) was introduced by @Gelman1997 and refined by @Park2004. It proceeds in two steps:

 1. Estimate a multilevel regression, predicting opinion using observed individual-level and unit-level covariates.
 2. Poststratify by taking the mean of each group's prediction weighted by their frequency in the subnational unit.

We can estimate the first-stage regression using the `lme4` package.

```{r multilevel regression, message=FALSE}
library(lme4)

model1 <- lmer(y ~ (1|x1) + (1|x2) + unit_covariate + (1|unit), data = trainset)
```

For the second stage, we need a **poststratification frame**. For the `SRP` package, it should come in the following format.

```{r PSFrame, message=FALSE}
PSFrame <- SRP::vignettePSFrame

PSFrame
```

Each row reports the empirical frequency for each unique combination of individual-level characteristics, repeated for each subnational unit. For example, the first row reports that there are 5482 individuals with $x_1 = 1$ and $x_2 = 1$ in Unit 1. 

Once we have both pieces of information -- the predictions and the frequencies -- poststratification simply requires taking a weighted average, using the `poststratify` function. Note that this function requires your poststratification frame to have two particular variables:

 * `unit`: the identity of the subnational unit.
 * `freq`: the empirical frequency for each cell.

```{r poststratification1, message=FALSE}
pred <- predict(model1, PSFrame, allow.new.levels = T)

mrp_estimates <- poststratify(pred, PSFrame)

mrp_estimates
```


## SRP

Stacked regression and poststratification (SRP) proceeds in the same fashion as MRP, but the first-stage predictions come from an ensemble model average generated through stacking. See @Ornstein2019 for technical details. 

To start, we must tune and estimate each of the component models separately. The following code estimates a hierarchical linear regression model, LASSO, random forest, KNN, and gradient boosting. 

```{r machine learning, message=FALSE}
library(glmnet)
library(ranger)
library(kknn)
library(xgboost)
library(caret)


#Estimate HLM
hlmFormula <- y ~ (1|x1) +  (1|x2) + unit_covariate + (1|unit) 
hlmModel <- lmer(hlmFormula, data = trainset)

#Tune LASSO
lasso_vars <- c("x1","x2","unit","unit_covariate")
lasso_factors <- c('x1', 'x2', 'unit') #which variables to convert to factors
trainset_lasso <- cleanDataLASSO(trainset, lasso_vars, lasso_factors)$trainset

lassoModel <- cv.glmnet(trainset_lasso, trainset$y, 
                        type.measure = "mse") 

#Tune KNN
knnFormula <- y ~ x1 + x2 + latitude + longitude + unit_covariate
knn_train <- train.kknn(knnFormula, data=trainset, kmax = 201) #Find best k (LOOCV)
k_best <- knn_train$best.parameters$k

#Tune Random Forest
forestFormula <- y ~ x1 + x2 + latitude + longitude + unit_covariate
forestModel <- ranger(forestFormula, data = trainset)

#Tune GBM
gbm_vars <- c("x1","x2","latitude","longitude","unit_covariate")
trainset_gbm <- cleanDataGBM(trainset=trainset, gbm_vars=gbm_vars)$trainset
#Create a custom 'xgb.DMatrix'. Faster computation
dtrain <- xgb.DMatrix(trainset_gbm, label = trainset$y)

#5-fold cross-validation; pick nrounds that minimizes RMSE
xgb.tune <- xgb.cv(data = dtrain, 
                   booster = "gbtree",
                   objective = "reg:squarederror",
                   eval_metric = "rmse",
                   eta = 0.02,
                   nrounds = 50 / 0.02, #Lower eta -> more trees
                   nfold = 5, 
                   verbose = F,
                   early_stopping_rounds = 20)

gbmModel <- xgboost(data = dtrain, 
                    booster = "gbtree",
                    objective = "reg:squarederror",
                    eval_metric = "rmse",
                    eta = 0.02,
                    verbose = F,
                    nrounds = xgb.tune$best_iteration)
```



Next, we will use the `getStackWeights()` function to estimate the optimal ensemble model average weights using 5-fold cross-validation.

```{r stacking, message=FALSE, warning=FALSE}
stackWeights <- getStackWeights(trainset = trainset,
                                hlmFormula = hlmFormula,
                                lasso_vars = lasso_vars,
                                lasso_factors = lasso_factors,
                                forestFormula = forestFormula,
                                knnFormula = knnFormula, k_best = k_best,
                                gbm_vars = gbm_vars, gbm_factors = NULL, 
                                gbm_params = list(eta = 0.02), gbm_tune = xgb.tune, 
                                nfolds = 5)

stackWeights %>% round(3)
```

Then we can poststratify as before.


```{r poststratification2, message=FALSE}
PSFrame_lasso <- cleanDataLASSO(PSFrame, lasso_vars = lasso_vars, lasso_factors = lasso_factors,
                                new_vars_lasso = colnames(trainset_lasso))$trainset
PSFrame_gbm <- cleanDataGBM(PSFrame, gbm_vars = gbm_vars)$trainset

M1 <- predict(hlmModel, PSFrame, allow.new.levels = T)
M2 <- predict(lassoModel, newx = PSFrame_lasso, s = lassoModel$lambda.min)
M3 <- kknn(knnFormula, train = trainset, test = PSFrame, k = k_best)$fitted.values
M4 <- predict(forestModel, PSFrame)$predictions
M5 <- predict(gbmModel, PSFrame_gbm)
M <- cbind(M1,M2,M3,M4,M5) %>% as.matrix

pred <- M %*% stackWeights

#Poststratify
srp_estimates <- poststratify(pred, PSFrame)

head(srp_estimates)
```


## Results

Because the data came from a simulation, we also know the true unit-level means. Let's see how our estimates compare. 

```{r plots, echo=FALSE, message=FALSE}
true_means <- SRP::vignetteTruth

names(mrp_estimates) <- c('unit','mrp_estimate')
names(srp_estimates) <- c('unit','srp_estimate')

results <- true_means %>%
  left_join(disag_estimates, by = 'unit') %>%
  left_join(mrp_estimates, by = 'unit') %>%
  left_join(srp_estimates, by = 'unit')

ggplot(results) + geom_point(aes(x=disag_estimate, y=true_mean)) + 
  xlab('Disaggregated Estimate') + ylab('True Mean') + 
  geom_abline(intercept = 0, slope = 1) +
  theme_bw()

ggplot(results) + geom_point(aes(x=mrp_estimate, y=true_mean)) + 
  xlab('MRP Estimate') + ylab('True Mean') + 
  geom_abline(intercept = 0, slope = 1) +
  theme_bw()

ggplot(results) + geom_point(aes(x=srp_estimate, y=true_mean)) + 
  xlab('SRP Estimate') + ylab('True Mean') + 
  geom_abline(intercept = 0, slope = 1) +
  theme_bw()

```

## Synthetic Poststratification

What if you do not have the joint frequency distribution for all your predictor variables at the subnational level? @Leemann2016 propose a method that instead uses marginal frequency distributions called **synthetic poststratification**. The approach proceeds by multiplying the marginal probabilities to create a *synethtic* joint distribution, assuming that the predictor variables are statistically independent. 

Note that this is a strong assumption. However, if the first-stage model is additively-separable, then both classical and synthetic poststratification produce identical results (see Appendix A in @Ornstein2019 for the proof). This implies that Multilevel Regression and Synthetic Poststratification (MrsP) can produce *strictly* superior estimates when the first-stage model is linear-additive, because one can include more predictor variables. 

The `SRP` package provides a function that can generate synthetic poststratification frames, called `getSyntheticPSFrame()`. 

### Using the Function

Suppose you have two (non-synthetic) frequency distributions describing the same population.

```{r PSFrames, echo=TRUE, message=FALSE}
PSFrame1 <- SRP::race
PSFrame2 <- SRP::education

PSFrame1
PSFrame2
```

To get the synthetic joint distribution, just call `getSyntheticPSFrame()`. Note that the resulting output is consistent with the marginal frequency distributions; add up all the observations with race == 1 in and it should yield the same frequency from `PSFrame1`.

```{r getSynthetic, echo=TRUE, message=FALSE}
PSFrame <- getSyntheticPSFrame(PSFrame1, PSFrame2)

PSFrame
```

If you want to generate a synthetic poststratification frame from more than one marginal distribution, simply repeat the process.

```{r getSynthetic2, echo=TRUE, message=FALSE}
PSFrame3 <- SRP::sex

PSFrame3

PSFrame <- getSyntheticPSFrame(PSFrame, PSFrame3)

PSFrame
```

## References
